### 2019-12-05 ch7_incgraph/content.tex
fix fontsize for page layout
~

%% FIGURE 2
\picfig{fig/figure2.pdf}{\LARGEfigure}{fig_efficient}{{\bf Several strategies are employed in order to reduce time and memory consumption.}
(A) Only the depth 3 neighbourhood of each modified edge needs to be explored in order to have visited all modified five-node graphlets. (B) A lookup table can be used to easily look up the graphlet index of a subgraph, by weighing each edge in a 5-node subgraph by a power of~2. (C) The same combination of five nodes is never repeated, as to avoid counting the same graphlet multiple times. (D) The adjacency matrix of the network is compressed in order to reduce memory usage.}

@@

%% FIGURE 4
\picfig{fig/figure4.pdf}{\hugefigure}{fig_method_reranking}{{\bf Predicted gene regulatory networks of model organisms are optimised to reduce the false positive rate.}
	A)~The number of redundant edges in each graphlet are counted. B) The network is optimised in order to obtain a lower redundancy over time. Two networks are shown, one before and one after the optimisation procedure. Edges coloured in red have been removed from the network after optimisation, green edges have been added. C) Starting from an empty network, the interactions are modified by iteratively evaluating the increase in redundancy of the next $k$ interactions, and adding the first edge for which its redundancy is less than the $90^\textrm{th}$ percentile redundancy.}

@@

\section{Supporting information}

@@

\begin{figure}[htb!]
	\begin{algorithmic}
		\Function{CalculateDelta}{Network $G'$, Node $n_0$, Node $n_1$}
		\State \Comment{Assuming an edge $(n_0, n_1)$ has just been added or removed from $G'$}
		\State $\Delta^- \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the removed graphlets}
		\State $\Delta^+ \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the added graphlets}
		\State $B_0 = \{n_0, n_1\}$ \Comment{A blacklist of nodes not to visit anymore}
		\State $e = (n_0, n_1)$ \Comment{Different name for the edge}
		\If{$e \in N'$}
		\State $(m^-, m^+) = (0, 1)$ \Comment{In other words, $m^- = 1$ iff $e \in N$ and $m^+ = 1$ iff $e \in N'$}
		\Else
		\State $(m^-, m^+) = (1, 0)$
		\EndIf
		\State $x_0 = 0$
		\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1), x_0, m^-, m^+}$ \Comment{Update delta matrices for current nodes}

@@

\begin{figure}[htb!]	
	\begin{algorithmic}
		\Function{BarabasiAlbert}{number of nodes $n$, degree $d$, \par\hspace*{3.72cm} number of operations $o$, offset exponent $x=1$}
		\State $D \gets \Call{Rep}{0, n}$ \Comment{All nodes start with 0 degree}
		\State $T \gets \Call{Rep}{\{\}, n}$ \Comment{List of targets of each node}
		\State $S \gets \Call{Rep}{\{\}, n}$ \Comment{List of sources of each node}
		\State $O \gets (\,)$ \Comment{Variable to store the generated operations in}

@@

\begin{figure}[htb!]
	\begin{algorithmic}
		\Function{ErdosRenyi}{number of nodes $n$, number of edges $e$, number of operations $o$}
		\State $P \gets \{(i, j) \,|\, i \in \{1 \,..\, n-1\} \wedge j \in \{0 \,..\, i-1\}\}$ \Comment{All possible interactions}
		\State $N_0 \gets$ Sample $e$ edges from $P$
		\State $N_c \gets N_0$
		\While{$|O|<o$}
		\State $e_a \gets $ Sample 1 edge from $P \setminus N_c$
		\State $e_r \gets $ Sample 1 edge from $N_c$
		\State $N_c \gets (N_c \setminus e_r) \cup \{e_a\}$
		\State $O \gets O + ((\textrm{ADD}, e_a), (\textrm{REM}, e_r))$
		\EndWhile
		\State $O \gets \Call{Head}{O, o}$ 
		\State \Return $(N_0, O)$ \Comment{Return the initial geometric network and $o$ operations}
		\EndFunction
	\end{algorithmic}
	\caption{
		\textbf{Pseudo code for generating an evolving Erd\H{o}s--R\'enyi (ER) network.} It first generates an ER network, and then generates $o$ operations such that at any time point, the network is or very closely resembles an ER network.
	}
	\label{code_er}
\end{figure}

\begin{figure}[htb!]
	\begin{algorithmic}
		\Function{Geometric}{number of nodes $n$, number of edges $e$, \par\hspace*{2.77cm} number of operations $o$, number of dimensions $d=3$}
		\State $P \gets$ Sample $n$ points from a multivariate continuous uniform distribution $U_{d}((0,1)^{d})$
		\State $D \gets$ Calculate distance matrix from $P$
		\State $N_0 \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{Initial network}
		\State $N_p \gets N_0$
		\State $O \gets (\,)$ \Comment{Variable to store the generated operations in}
		\While{$|O| < o$} \Comment{Modify network until $O$ is sufficiently large}
		\State $i \gets $ Sample 1 index from $\{0\,..\,n-1\}$
		\State $P[i] \gets$ Sample 1 point from a $U_{d}((0,1)^{d})$ \Comment{Give node $i$ a new location}
		\State $D[i,] \gets D[,i] \gets $ Calculate new distances between node $i$ and all other nodes
		\State $N_c \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{New network}
		\State $O_a \gets \{(\textrm{ADD}, e) \,|\, e \not\in E(N_p) \wedge e \in E(N_c) \}$ \Comment{Gather added edges}
		\State $O_r \gets \{(\textrm{REM}, e) \,|\, e \in E(N_p) \wedge e \not\in E(N_c) \}$ \Comment{Gather removed edges}
		\State $O \gets O + \Call{Shuffle}{O_a + O_r}$ \Comment{Append new operations to $O$}
		\State $N_p \gets N_c$
		\EndWhile
		\State $O \gets \Call{Head}{O, o}$ 
		\State \Return $(N_0, O)$ \Comment{Return the initial geometric network and $o$ operations}
		\EndFunction
	\end{algorithmic}
	\caption{
		\textbf{Pseudo code for generating an evolving geometric network.} It first generates a geometric network, and then generates $o$ operations such that at any time point, the network is or very closely resembles a geometric network.
	}
	\label{code_geo}
\end{figure}

>>>>

%% FIGURE 2
\picfig{fig/figure2.pdf}{.7\linewidth}{fig_efficient}{{\bf Several strategies are employed in order to reduce time and memory consumption.}
(A) Only the depth 3 neighbourhood of each modified edge needs to be explored in order to have visited all modified five-node graphlets. (B) A lookup table can be used to easily look up the graphlet index of a subgraph, by weighing each edge in a 5-node subgraph by a power of~2. (C) The same combination of five nodes is never repeated, as to avoid counting the same graphlet multiple times. (D) The adjacency matrix of the network is compressed in order to reduce memory usage.}

@@

%% FIGURE 4
\picfig{fig/figure4.pdf}{.9\linewidth}{fig_method_reranking}{{\bf Predicted gene regulatory networks of model organisms are optimised to reduce the false positive rate.}
	A)~The number of redundant edges in each graphlet are counted. B) The network is optimised in order to obtain a lower redundancy over time. Two networks are shown, one before and one after the optimisation procedure. Edges coloured in red have been removed from the network after optimisation, green edges have been added. C) Starting from an empty network, the interactions are modified by iteratively evaluating the increase in redundancy of the next $k$ interactions, and adding the first edge for which its redundancy is less than the $90^\textrm{th}$ percentile redundancy.}

@@

\section{Supplemental information}

@@

\begin{figure}[htb!]
	\small
	\begin{algorithmic}
		\Function{CalculateDelta}{Network $G'$, Node $n_0$, Node $n_1$}
		\State \Comment{Assuming an edge $(n_0, n_1)$ has just been added or removed from $G'$}
		\State $\Delta^- \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the removed graphlets}
		\State $\Delta^+ \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the added graphlets}
		\State $B_0 = \{n_0, n_1\}$ \Comment{A blacklist of nodes not to visit anymore}
		\State $e = (n_0, n_1)$ \Comment{Different name for the edge}
		\If{$e \in N'$}
		\State $(m^-, m^+) = (0, 1)$ \Comment{In other words, $m^- = 1$ iff $e \in N$ and $m^+ = 1$ iff $e \in N'$}
		\Else
		\State $(m^-, m^+) = (1, 0)$
		\EndIf
		\State $x_0 = 0$
		\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1), x_0, m^-, m^+}$ \Comment{Update delta matrices for current nodes}

@@

\begin{figure}[htb!]	
	\small
	\begin{algorithmic}
		\Function{BarabasiAlbert}{number of nodes $n$, degree $d$, \par\hspace*{3.72cm} number of operations $o$, offset exponent $x=1$}
		\State $D \gets \Call{Rep}{0, n}$ \Comment{All nodes start with 0 degree}
		\State $T \gets \Call{Rep}{\{\}, n}$ \Comment{List of targets of each node}
		\State $S \gets \Call{Rep}{\{\}, n}$ \Comment{List of sources of each node}
		\State $O \gets (\,)$ \Comment{Variable to store the generated operations in}

@@

\begin{figure}[htb!]
	\small
	\begin{algorithmic}
		\Function{ErdosRenyi}{number of nodes $n$, number of edges $e$, number of operations $o$}
		\State $P \gets \{(i, j) \,|\, i \in \{1 \,..\, n-1\} \wedge j \in \{0 \,..\, i-1\}\}$ \Comment{All possible interactions}
		\State $N_0 \gets$ Sample $e$ edges from $P$
		\State $N_c \gets N_0$
		\While{$|O|<o$}
		\State $e_a \gets $ Sample 1 edge from $P \setminus N_c$
		\State $e_r \gets $ Sample 1 edge from $N_c$
		\State $N_c \gets (N_c \setminus e_r) \cup \{e_a\}$
		\State $O \gets O + ((\textrm{ADD}, e_a), (\textrm{REM}, e_r))$
		\EndWhile
		\State $O \gets \Call{Head}{O, o}$ 
		\State \Return $(N_0, O)$ \Comment{Return the initial geometric network and $o$ operations}
		\EndFunction
	\end{algorithmic}
	\caption{
		\textbf{Pseudo code for generating an evolving Erd\H{o}s--R\'enyi (ER) network.} It first generates an ER network, and then generates $o$ operations such that at any time point, the network is or very closely resembles an ER network.
	}
	\label{code_er}
\end{figure}

\begin{figure}[htb!]
	\small
	\begin{algorithmic}
		\Function{Geometric}{number of nodes $n$, number of edges $e$, \par\hspace*{2.77cm} number of operations $o$, number of dimensions $d=3$}
		\State $P \gets$ Sample $n$ points from a multivariate continuous uniform distribution $U_{d}((0,1)^{d})$
		\State $D \gets$ Calculate distance matrix from $P$
		\State $N_0 \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{Initial network}
		\State $N_p \gets N_0$
		\State $O \gets (\,)$ \Comment{Variable to store the generated operations in}
		\While{$|O| < o$} \Comment{Modify network until $O$ is sufficiently large}
		\State $i \gets $ Sample 1 index from $\{0\,..\,n-1\}$
		\State $P[i] \gets$ Sample 1 point from a $U_{d}((0,1)^{d})$ \Comment{Give node $i$ a new location}
		\State $D[i,] \gets D[,i] \gets $ Calculate new distances between node $i$ and all other nodes
		\State $N_c \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{New network}
		\State $O_a \gets \{(\textrm{ADD}, e) \,|\, e \not\in E(N_p) \wedge e \in E(N_c) \}$ \Comment{Gather added edges}
		\State $O_r \gets \{(\textrm{REM}, e) \,|\, e \in E(N_p) \wedge e \not\in E(N_c) \}$ \Comment{Gather removed edges}
		\State $O \gets O + \Call{Shuffle}{O_a + O_r}$ \Comment{Append new operations to $O$}
		\State $N_p \gets N_c$
		\EndWhile
		\State $O \gets \Call{Head}{O, o}$ 
		\State \Return $(N_0, O)$ \Comment{Return the initial geometric network and $o$ operations}
		\EndFunction
	\end{algorithmic}
	\caption{
		\textbf{Pseudo code for generating an evolving geometric network.} It first generates a geometric network, and then generates $o$ operations such that at any time point, the network is or very closely resembles a geometric network.
	}
	\label{code_geo}
\end{figure}
